---
title: "What Metrics Matter?"
subtitle: "A Public Equity Investor's Perspective"
author: "TiE Denver Business for Breakfast"

date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(tidyverse)
public_vs_private <- tibble(Similarities = c("fiduciary duty", "impact of macroeconomic conditions", 
                                 "impact of capital market conditions", "impact of investor risk appetite",
                                 "same competitive environment", "importance of expectations"), 
                Differences = c("liquidity", "data availability", "reporting requirements", 
                                "investor universe", "investor motivations and constraints",
                                "relationship between management and investors"))

porter <- tibble(Dimension = c("threat of new entrants", "threat of substitutes",
                               "bargaining power of customers", "bargaining power of suppliers",
                               "competitive rivalry"),
                 High = c("food truck", "cable TV", "AAPL suppliers", "oilfield well integrity services", "ride sharing"),
                 Low = c("regulated utility", "cloud computing", "Lululemon", "Amazon", "oncology drug development"))

vc_playbook <- tibble(Characteristic = c("relevant metrics", "customer focus", "vision", "agility", "attitude"),
                      Comments = c("is management focused on a small, concrete list of objectives?",
                                   "does management understand the customer deeply?",
                                   "does the CEO have high expectations and a compelling vision?",
                                   "is the company willing and able to change direction quickly if necessary?",
                                   "does the company value curiousity and experimentation?"))

valuation <- tibble(Metric = c("trailing P/E", "forward P/E", "price to sales", "EV/EBITDA", 
                               "free cash flow yield", "dividend yield"),
                    Comments = c("popular in the financial press", "useful for assessing duration expectations", "useful for high growth companies",
                                 "accounts for capital structure", "important long term, but places emphasis on past performance",
                                 "useful for capital intensive companies with low growth expectations"))

quality <- tibble(Metric = c("ROIC", "earnings information ratio", "incremental ROIC",
                             "debt to equity", "operating leverage", "cash conversion cycle"),
                  Comments = c("what are you doing with what you have?", "how consistent are your profits?",
                               "what can you do with an additional unit of investment?", "is debt service an issue or is WACC too high?",
                               "how does top line growth impact margins?", "do you have leverage over your suppliers and customers?"))

timing <- tibble(Metric = c("12 month trailing return", "earnings revisions", "margin expansion",
                            "estimates dispersion", "growth inflection", "relative valuation"),
                  Comments = c("commonly used momentum factor", "good indicator of changing expectations",
                               "operational momentum indicator", "divergent expections can offer opportunities",
                               "this is great, but the new growth rate must persist", "does the asset look cheap vs. history on a relative basis?"))

capital_deployment <- tibble(Metric = c("dividend yield", "share buybacks", "management compensation",
                                        "capex vs. depreciation", "acquisition terms", "reinvestment rate"),
                             Comments = c("is the company returning the capital it can't invest above it's WACC?",
                                          "is the company buying back shares? if so, is the valuation low or high?",
                                          "is there a large option pool for management?", "is the company reinvesting adequately to sustain current revenues?",
                                          "cash or stock? what are the synergy expectations?", "what proportion of the company's profits are being reinvested?"))

esg <- tibble(Category = c("Governance", "Social", "Governance", "Environmental", "Governance", "Social"),
              Example = c("Is the CEO's compensation in line with industry peers?",
                          "Is the company aware of how its products and operations affect the communities in which it operates?",
                          "Are the majority of board members independent?",
                          "Does the company have a reputation of careless or short sighted operating practices?",
                          "How are the voting rights distributed? Does the CEO have voting control?",
                          "Does the company make an effort to develop a diverse workforce?"))
```


# About Me

- Chris Lundberg

- Research Associate at [Segall Bryant & Hamill](https://www.sbhic.com)

- responsible for equity risk and alpha modeling, asset allocation, and alternative manager selection 

- CFA charterholder

- email: clundberg@sbhic.com

- github repo for slides: 
  
    https://github.com/chris-lundberg/measure-what-matters

---

# Public vs. Private Markets

```{r, echo=FALSE}
knitr::kable(public_vs_private, format = 'html')
```

---
# Investment Decision Making

## Three Paradigms

* Qualitative

* Quantitative

* ESG
---
class: inverse, middle, center

# Qualitative: Competition and Execution
---
# Porter's Five Forces

```{r, echo=FALSE}
knitr::kable(porter, format = 'html')
```
---
# A Culture of Success

```{r, echo=FALSE}
knitr::kable(vc_playbook, format = 'html')
```

.footnote[\*borrowed heavily from *The Lean Startup* and *Measure What Matters*]
---
class: inverse, middle, center

# Quantitative: Return Predictive Signals
---
# Common Metrics: Valuation

```{r, echo=FALSE}
knitr::kable(valuation, format = 'html')
```
---
# Common Metrics: Quality

```{r, echo=FALSE}
knitr::kable(quality, format = 'html')
```
---
# Common Metrics: Timing

```{r, echo=FALSE}
knitr::kable(timing, format = 'html')
```
---
# Common Metrics: Capital Deployment

```{r, echo=FALSE}
knitr::kable(capital_deployment, format = 'html')
```
---
class: inverse, middle, center

# ESG: Environmental, Social, Governance
---
# Principles of Responsible Investing

```{r, echo=FALSE}
knitr::kable(esg, format = 'html')
```
---

# References

[*Amazon Shareholder Letters* - Jeff Bezos](https://ir.aboutamazon.com/annual-reports)

[*The Innovator's Dilemma* - Clayton Christensen](https://www.amazon.com/Innovators-Dilemma-Technologies-Management-Innovation/dp/1633691780/ref=sr_1_1?crid=2OC8OAB127CEK&keywords=the+innovator%27s+dilemma&qid=1581053880&sprefix=the+innovat%2Caps%2C199&sr=8-1)

[*Measure What Matters* - John Doerr](https://www.amazon.com/Measure-What-Matters-Google-Foundation/dp/0525536221/ref=sr_1_1?crid=BGL8P7LSPF3D&keywords=measure+what+matters+by+john+doerr&qid=1581052332&sprefix=measure+what+ma%2Caps%2C186&sr=8-1)

[*No Mercy/No Malice* - Scott Galloway's blog](https://www.profgalloway.com/)

[*The Supraview of Return Predictive Signals* - Green et al.](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2062464)

[*The New Tycoons* - Jason Kelly](https://www.amazon.com/New-Tycoons-Trillion-Industry-Everything/dp/1118205464/ref=tmm_hrd_swatch_0?_encoding=UTF8&qid=1581055394&sr=8-1)

[*One Up on Wall Street* - Peter Lynch](https://www.amazon.com/One-Up-Wall-Street-Already/dp/0743200403/ref=sr_1_1?crid=3DRGKOSJZBYBD&keywords=one+up+on+wall+street+peter+lynch&qid=1581055441&sprefix=one+up+on%2Caps%2C533&sr=8-1)

[*Expectations Investing* - Michael Mauboussin](https://www.amazon.com/Expectations-Investing-Reading-Prices-Returns/dp/159139127X/ref=sr_1_1?crid=29OCUUWCS0EQT&keywords=expectations+investing&qid=1581052396&sprefix=expectations+inv%2Caps%2C195&sr=8-1)

[*Offense is the New Defense* - Partners Group 2020 Outlook](https://www.partnersgroup.com/en/news-views/corporate-news/current/detail/article/partners-group-publishes-private-markets-outlook-for-2020-offense-is-the-new-defense/)

[*Competitive Strategy* - Michael Porter](https://www.amazon.com/Competitive-Strategy-Techniques-Industries-Competitors/dp/0684841487/ref=sr_1_2?crid=FLLJR72XGUWY&keywords=porter%27s+five+forces&qid=1581052133&sprefix=porter%27s+five%2Caps%2C183&sr=8-2)

[*The Lean Startup* - Eric Ries](https://www.amazon.com/Lean-Startup-Entrepreneurs-Continuous-Innovation/dp/0307887898/ref=sr_1_1?keywords=lean+startup&qid=1581053834&sr=8-1)

[United Nations *Principles of Responsible Investing*](https://www.unpri.org/)
